---
title: "Pull data"
author: "Abby Lewis"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("remotes")
#remotes::install_github("eco4cast/neon4cast")
library(neon4cast)
library(tidyverse)
library(lubridate)
install.packages("rMR")
library(rMR)
```

Neon4cast has functions for a lot of what we are hoping to do! Here I am reviving a deprecated function from that package to download data for a given target (why was this function deprecated?? unsure)
```{r}
#Using this deprecated function from neon4cast
#https://github.com/eco4cast/neon4cast/blob/main/R/download_target.R
download_target <- function(theme = c("aquatics", "beetles",
                                      "phenology", "terrestrial_30min",
                                      "terrestrial_daily","ticks")){  
    theme <- match.arg(theme)
    
    target_file <- switch(theme,
                          aquatics = "aquatics-targets.csv.gz",
                          beetles = "beetles-targets.csv.gz",
                          phenology = "phenology-targets.csv.gz",
                          terrestrial_daily = "terrestrial_daily-targets.csv.gz",
                          terrestrial_30min = "terrestrial_30min-targets.csv.gz",
                          ticks = "ticks-targets.csv.gz"
    )
    download_url <- paste0("https://data.ecoforecast.org/targets/",
                           theme, "/", target_file)
  
  readr::read_csv(download_url, show_col_types = FALSE,
                  lazy = FALSE, progress = FALSE)%>% 
  as_tibble(index=time, key=siteID)
  
}
```

Using a neon4cast example found here: https://github.com/eco4cast/neon4cast-example
One thing that is nice about this example is we can easily set it up to run automatically using code from this repository
```{r}
dir.create("drivers", showWarnings = FALSE)

forecast_date <- Sys.Date()
noaa_date <- Sys.Date() - days(1)  #Need to use yesterday's NOAA forecast because today's is not available yet

#Step 0: Define team name, team members, and theme

team_name <- "EFI Theory"

team_list <- list(list(individualName = list(givenName = "Abby", 
                                             surName = "Lewis"),
                       organizationName = "Virginia Tech",
                       electronicMailAddress = "aslewis@vt.edu"),
                  list(individualName = list(givenName = "Sikander", 
                                             surName = "Khare"),
                       organizationName = "University of Florida",
                       electronicMailAddress = " "))

theme = c("terrestrial_30min") #Define theme. Options: aquatics, beetles, phenology, terrestrial_30min, terrestrial_daily, ticks

#Step 1: Download latest target data and site description data

target = download_target(theme)
type = ifelse(theme%in% c("terrestrial_30min", "terrestrial_daily"),"terrestrial",theme)

site_data <- readr::read_csv("https://raw.githubusercontent.com/eco4cast/neon4cast-targets/main/NEON_Field_Site_Metadata_20220412.csv") %>%
  filter(get(type)==1)

#Step 2: Get drivers (currently noaa forecasts aren't working)
sites <- unique(target$site_id)

```

Separating the code below because noaa forecast downloads aren't working
```{r}
df_past <- neon4cast::noaa_stage3()#This should be a function from neon4cast but looks like something got messed up. Will need to fix later

df_future <- neon4cast::noaa_stage2() #Same as above

noaa_past <- df_past |> 
  dplyr::filter(site_id %in% sites,
                variable == "air_temperature") |> 
  dplyr::collect()

noaa_future <- df_future |> 
  dplyr::filter(cycle == 0,
                start_date == as.character(noaa_date),
                time >= lubridate::as_datetime(forecast_date), 
                variable == "air_temperature") |> 
  dplyr::collect()

# Step 2.4 Aggregate (to day) and convert units of drivers

noaa_past_mean <- noaa_past %>% 
  mutate(date = as_date(time)) %>% 
  group_by(date, site_id) %>% 
  summarize(air_temperature = mean(predicted, na.rm = TRUE), .groups = "drop") %>% 
  rename(time = date) %>% 
  mutate(air_temperature = air_temperature - 273.15)


noaa_future <- noaa_future %>% 
  mutate(time = as_date(time)) %>% 
  group_by(time, site_id, ensemble) |> 
  summarize(air_temperature = mean(predicted), .groups = "drop") |> 
  mutate(air_temperature = air_temperature - 273.15) |> 
  select(time, site_id, air_temperature, ensemble)

#Step 2.5: Merge in past NOAA data into the targets file, matching by date.
target <- target |> 
  select(time, site_id, variable, observed) |> 
  filter(variable %in% c("temperature", "oxygen")) |> 
  pivot_wider(names_from = "variable", values_from = "observed")

target <- left_join(target, noaa_past_mean, by = c("time","site_id"))

ggplot(target, aes(x = temperature, y = air_temperature)) +
  geom_point() +
  labs(x = "NEON water temperature (C)", y = "NOAA air temperature (C)") +
  facet_wrap(~site_id)

```

Sikander- this is probably the relevant part for you to modify!
```{r}
#Step 3.0: Generate forecasts for each site

forecast <- NULL

for(i in 1:length(sites)){
  
  # Get site information for elevation
  site_info <- site_data %>% filter(field_site_id == sites[i]) 
  
  site_target <- target |> 
    filter(site_id == sites[i])
  
  noaa_future_site <- noaa_future |> 
    filter(site_id == sites[i])
  
  if(length(which(!is.na(site_target$air_temperature) & !is.na(site_target$temperature))) > 0){
    
    #Fit linear model based on past data: water temperature = m * air temperature + b
    fit <- lm(site_target$temperature~site_target$air_temperature)
    
    #use linear regression to forecast water temperature for each ensemble member
    forecasted_temperature <- fit$coefficients[1] + fit$coefficients[2] * noaa_future_site$air_temperature
    
    #use forecasted temperature to predict oyxgen by assuming that oxygen is saturated.  
    forecasted_oxygen <- rMR::Eq.Ox.conc(forecasted_temperature, elevation.m = ,site_info$field_mean_elevation_m, 
                                         bar.press = NULL, 
                                         bar.units = NULL,
                                         out.DO.meas = "mg/L",
                                         salinity = 0, 
                                         salinity.units = "pp.thou")
    
    temperature <- tibble(time = noaa_future_site$time,
                          site_id = sites[i],
                          ensemble = noaa_future_site$ensemble,
                          predicted = forecasted_temperature,
                          variable = "temperature")
    
    oxygen <- tibble(time = noaa_future_site$time,
                     site_id = sites[i],
                     ensemble = noaa_future_site$ensemble,
                     predicted = forecasted_oxygen,
                     variable = "oxygen")
    
    
    #Build site level dataframe.  Note we are not forecasting chla
    forecast <- dplyr::bind_rows(forecast, temperature, oxygen)
  }
}

forecast <- forecast |> 
  mutate(start_time = forecast_date) |> #start_time is today
  select(time, start_time, site_id, variable, ensemble, predicted)

#Visualize forecast.  Is it reasonable?
forecast %>% 
  ggplot(aes(x = time, y = predicted, group = ensemble)) +
  geom_line() +
  facet_grid(variable~site_id, scale ="free")

#Forecast output file name in standards requires for Challenge.  
# csv.gz means that it will be compressed
forecast_file <- paste0("aquatics","-",min(forecast$time),"-",team_name,".csv.gz")

#Write csv to disk
write_csv(forecast, forecast_file)

#Confirm that output file meets standard for Challenge
#neon4cast::forecast_output_validator(forecast_file)
```

Will need to change this later
```{r}
# Step 4: Generate metadata

model_metadata = list(
  forecast = list(
    model_description = list(
      forecast_model_id =  "air2waterSat",  #What goes here
      name = "Air temperatuer to water temperature linear regression plus assume saturated oxygen", 
      type = "empirical",  
      repository = "https://github.com/rqthomas/neon4cast-example" 
    ),
    initial_conditions = list(
      status = "absent"
    ),
    drivers = list(
      status = "propagates",
      complexity = 1, #Just temperature
      propagation = list( 
        type = "ensemble", 
        size = 31) 
    ),
    parameters = list(
      status = "absent"
    ),
    random_effects = list(
      status = "absent"
    ),
    process_error = list(
      status = "absent"
    ),
    obs_error = list(
      status = "absent"
    )
  )
)

metadata_file <- neon4cast::generate_metadata(forecast_file, team_list, model_metadata)

# Step 5: Submit forecast!


neon4cast::submit(forecast_file = forecast_file, metadata = metadata_file, ask = FALSE)
```
