---
title: "Pull data"
author: "Abby Lewis"
date: "2022-09-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("remotes")
#remotes::install_github("eco4cast/neon4cast")
library(neon4cast)
library(tidyverse)
library(lubridate)
library(tsibble)
library(fable)
#install.packages("rMR")
library(rMR)
```


TO DO:
set timestamp for each challenge
format random walk to avoid negative values



Neon4cast has functions for a lot of what we are hoping to do! Here I am reviving a deprecated function from that package to download data for a given target (why was this function deprecated?? unsure)
```{r}
#Using this deprecated function from neon4cast
#https://github.com/eco4cast/neon4cast/blob/main/R/download_target.R
download_target <- function(theme = c("aquatics", "beetles",
                                      "phenology", "terrestrial_30min",
                                      "terrestrial_daily","ticks")){  
    theme <- match.arg(theme)
    
    target_file <- switch(theme,
                          aquatics = "aquatics-targets.csv.gz",
                          beetles = "beetles-targets.csv.gz",
                          phenology = "phenology-targets.csv.gz",
                          terrestrial_daily = "terrestrial_daily-targets.csv.gz",
                          terrestrial_30min = "terrestrial_30min-targets.csv.gz",
                          ticks = "ticks-targets.csv.gz"
    )
    download_url <- paste0("https://data.ecoforecast.org/targets/",
                           theme, "/", target_file)
  
  readr::read_csv(download_url, show_col_types = FALSE,
                  lazy = FALSE, progress = FALSE)#%>% 
  #as_tibble(index=time, key=siteID)
  
}
```


Using a neon4cast example found here: https://github.com/eco4cast/neon4cast-example
One thing that is nice about this example is we can easily set it up to run automatically using code from Quinn

```{r}
dir.create("drivers", showWarnings = FALSE)

forecast_date <- Sys.Date()
noaa_date <- Sys.Date() - days(1)  #Need to use yesterday's NOAA forecast because today's is not available yet

#Step 0: Define team name, team members, and theme

team_name <- "EFI Theory"

team_list <- list(list(individualName = list(givenName = "Abby", 
                                             surName = "Lewis"),
                       organizationName = "Virginia Tech",
                       electronicMailAddress = "aslewis@vt.edu")
                  )

model_name = "Test"
model_id = "theory_test"

theme = c("beetles") #Define theme. Options: aquatics, beetles, phenology, terrestrial_30min, terrestrial_daily, ticks

#Step 1: Download latest target data and site description data

target = download_target(theme)
type = ifelse(theme%in% c("terrestrial_30min", "terrestrial_daily"),"terrestrial",theme)

site_data <- readr::read_csv("https://raw.githubusercontent.com/eco4cast/neon4cast-targets/main/NEON_Field_Site_Metadata_20220412.csv") %>%
  filter(get(type)==1)

#Step 2: Get drivers (currently noaa forecasts aren't working)
if("site_id" %in% colnames(target)){
  target = target%>%
    rename(siteID = site_id)
}
sites <- unique(target$siteID)


#Step 3.0: Generate forecasts for each site

# Get site information for elevation
site_info <- site_data %>% filter(field_site_id == sites[i]) 

site_target <- target |> 
  filter(siteID %in% sites)

#noaa_future_site <- noaa_future |> 
#  filter(site_id == sites[i])

# drop last 35 days and use explicit NAs for gaps in timeseries
blinded_target <- site_target %>% 
  #filter(time < max(time) - 35) %>% 
  group_by(time, siteID)%>%
  summarize_all(mean)%>%
  as_tsibble(index=time, key=siteID)%>%
  fill_gaps()

if(type == "aquatics")           {vars = c("temperature","oxygen","chla")}
if(type == "ticks")              {vars = c("amblyomma_americanum")}
if(type == "phenology")          {vars = c("gcc_90","rcc_90")}
if(type == "beetles")            {vars = c("abundance","richness")}
if(type == "terrestrial_daily")  {vars = c("nee","le")}

forecast <- data.frame(time=as.Date(integer(0),origin = "1970-01-01"),
                start_time=as.Date(integer(0),origin = "1970-01-01"),
                siteID=character(),
                variable=character(),
                ensemble=numeric(),
                predicted=numeric())
for(var in vars){
  
  # A simple random walk forecast, see ?fable::RW
  var_fc <- blinded_target%>%
    model(null = RW(!!sym(var))) %>%
    forecast(h = "35 days") %>%
    efi_format()%>%
    pivot_longer(cols = all_of(var), names_to = "variable", values_to = "predicted")
  
  # combine into single table
  forecast <- forecast%>%
    full_join(var_fc)
  
}

## Write the forecast to a file following EFI naming conventions:
forecast_file <- glue::glue("{theme}-{date}-{team}.csv.gz",
                            theme = theme, 
                            date=Sys.Date(),
                            team = team_name)
write_csv(forecast, forecast_file)

forecast <- forecast |> 
  mutate(start_time = forecast_date) |> #start_time is today
  select(time, start_time, siteID, variable, ensemble, predicted)

#Visualize forecast.  Is it reasonable?
forecast %>% 
  ggplot(aes(x = time, y = predicted, group = ensemble)) +
  geom_line() +
  facet_grid(variable~siteID, scale ="free")

#Forecast output file name in standards requires for Challenge.  
# csv.gz means that it will be compressed
forecast_file <- paste0(theme,"-",min(forecast$time),"-",team_name,".csv.gz")

#Write csv to disk
write_csv(forecast, forecast_file)

#Confirm that output file meets standard for Challenge
neon4cast::forecast_output_validator(forecast_file)

```


Submit!

```{r}
# Step 4: Generate metadata

model_metadata = list(
  forecast = list(
    model_description = list(
      forecast_model_id =  model_id, 
      name = model_name, 
      type = "empirical",  
      repository = "https://github.com/abbylewis/EFI_Theory" 
    ),
    initial_conditions = list(
      status = "absent"
    ),
    drivers = list(
      status = "propagates",
      complexity = 1, #Just temperature
      propagation = list( 
        type = "ensemble", 
        size = 31) 
    ),
    parameters = list(
      status = "absent"
    ),
    random_effects = list(
      status = "absent"
    ),
    process_error = list(
      status = "absent"
    ),
    obs_error = list(
      status = "absent"
    )
  )
)

metadata_file <- neon4cast::generate_metadata(forecast_file, team_list, model_metadata)

# Step 5: Submit forecast!


neon4cast::submit(forecast_file = forecast_file, metadata = metadata_file, ask = FALSE)
```





Consider adding back in later (not currently working)


Separating the code below because noaa forecast downloads weren't working (they are now!)
```{r}
df_past <- neon4cast::noaa_stage3()#This should be a function from neon4cast but looks like something got messed up. Will need to fix later
neon4cast::noaa_stage1()
df_future <- neon4cast::noaa_stage2() #Same as above

data = neon4cast::download_noaa(sites)
read.csv(data)
path = download_forecast("phenology")
list.files(paste0(path,"/downloaded_packages"))

noaa_past <- df_past |> 
  dplyr::filter(site_id %in% sites,
                variable == "air_temperature") |> 
  dplyr::collect()

noaa_future <- df_future |> 
  dplyr::filter(cycle == 0,
                start_date == as.character(noaa_date),
                time >= lubridate::as_datetime(forecast_date), 
                variable == "air_temperature") |> 
  dplyr::collect()

# Step 2.4 Aggregate (to day) and convert units of drivers

noaa_past_mean <- noaa_past %>% 
  mutate(date = as_date(time)) %>% 
  group_by(date, site_id) %>% 
  summarize(air_temperature = mean(predicted, na.rm = TRUE), .groups = "drop") %>% 
  rename(time = date) %>% 
  mutate(air_temperature = air_temperature - 273.15)


noaa_future <- noaa_future %>% 
  mutate(time = as_date(time)) %>% 
  group_by(time, site_id, ensemble) |> 
  summarize(air_temperature = mean(predicted), .groups = "drop") |> 
  mutate(air_temperature = air_temperature - 273.15) |> 
  select(time, site_id, air_temperature, ensemble)

#Step 2.5: Merge in past NOAA data into the targets file, matching by date.
target <- target |> 
  select(time, site_id, variable, observed) |> 
  filter(variable %in% c("temperature", "oxygen")) |> 
  pivot_wider(names_from = "variable", values_from = "observed")

target <- left_join(target, noaa_past_mean, by = c("time","site_id"))

ggplot(target, aes(x = temperature, y = air_temperature)) +
  geom_point() +
  labs(x = "NEON water temperature (C)", y = "NOAA air temperature (C)") +
  facet_wrap(~site_id)

```


